### 2021-05-12 report/paper.tex
Add template for group report

\documentclass[]{usiinfprospectus}

\captionsetup{labelfont={bf}}

\author{Bevilacqua Joey \\[3pt] De Vita Gianmarco  \\[3pt] Luini Alessandro  \\[3pt] Maggioni Claudio}

\title{\centerline{Sudoku}}
\versiondate{\today}

\abstract {
Abstract}

\begin{document}
\maketitle
\tableofcontents
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{introduction}

%%%%%
\bibliographystyle{abbrv}
\bibliography{references}

\end{document}

### 2021-05-12 solver/app.py
Initial implementation of the z3 solver

@@ -1,5 +1,48 @@
 from flask import Flask
+from z3 import And, Bool, Int, Or, solve
+
+
+def is_valid(*data: str):
+    x = []
+    for i in range(0, 9):
+        if data[i].isnumeric():
+            x[i] = int(data[i])
+        else:
+            x[i] = Int(data[i])
+
+    and_args = []
+    for d in range(0, 9):
+        or_args = []
+        for i in range(0, 9):
+            or_args.append(Bool(x[i] == d))
+        and_args.append(Or(*or_args))
+
+    return And(*and_args)
+
+
+def solve_sudoku(sudoku):
+    rows_check = []
+    for i in range(0, 9):
+        rows_check.append(is_valid(sudoku[i, 0], sudoku[i, 1], sudoku[i, 2],
+                                   sudoku[i, 3], sudoku[i, 4], sudoku[i, 5],
+                                   sudoku[i, 6], sudoku[i, 7], sudoku[i, 8]))
+    columns_check = []
+    for i in range(0, 9):
+        columns_check.append(is_valid(sudoku[0, i], sudoku[1, i], sudoku[2, i],
+                                      sudoku[3, i], sudoku[4, i], sudoku[5, i],
+                                      sudoku[6, i], sudoku[7, i], sudoku[8, i]))
+    squares_check = []
+    for i in range(0, 9, step=3):
+        for j in range(0, 9, step=3):
+            squares_check.append(is_valid(sudoku[i, j], sudoku[i, j + 1], sudoku[i, j + 2],
+                                          sudoku[i + 1, j], sudoku[i + 1, j + 1], sudoku[i + 1, j + 2],
+                                          sudoku[i + 2, j], sudoku[i + 2, j + 1], sudoku[i + 2, j + 2]))
+
+    return solve(And(And(*rows_check), And(*columns_check), And(*squares_check)))
+
+
 app = Flask(__name__)
+
 
 @app.route("/")
 def hello_world():

### 2021-05-12 report/paper.tex
Started writing stuff into the group report

@@
 
-\author{Bevilacqua Joey \\[3pt] De Vita Gianmarco  \\[3pt] Luini Alessandro  \\[3pt] Maggioni Claudio}
+\author{\centerline{Bevilacqua Joey} \\[3pt] \centerline{De Vita Gianmarco}  \\[3pt] \centerline{Luini Alessandro}  \\[3pt] \centerline{Maggioni Claudio}}
 
@@
 \versiondate{\today}
+\usepackage{boldline} 
+\usepackage{enumitem}
 
 \abstract {
-Abstract}
+The aim of this project is to make an Encoder and Decoder for a SAT-solver in order to decide whether a problem is satisfiable or unsatisfiable. WRITE A BETTER ABSTRACT
+}
+
+% For Sudoku drawing
+\usepackage{tikz}
+\newcounter{row}
+\newcounter{col}
+
+\newcommand\setrow[9]{
+  \setcounter{col}{1}
+  \foreach \n in {#1, #2, #3, #4, #5, #6, #7, #8, #9} {
+    \edef\x{\value{col} - 0.5}
+    \edef\y{9.5 - \value{row}}
+    \node[anchor=center] at (\x, \y) {\n};
+    \stepcounter{col}
+  }
+  \stepcounter{row}
+}
 
@@
 \section{Introduction} \label{introduction}
+\section{Problem definition} \label{problem}
+Sudokus are puzzles based upon number placement.
+They are classically structured in a $9\times 9$ grid with 81 cells, divided into $3\times 3$ boxes.
+In order to be satisfiable, the following statements must hold:
+\begin{itemize}%[label={(\arabic*)}]
+\item In every cell there is exactly one number $c \in \{1, 2, ..., 9 \}$
+\item No number may appear more than once in the same row or column.
+\item No number may appear more than once in the same box.
+\end{itemize}
+Often Sudoku puzzles are initialized with some values placed.
 
+%Draw Sudoku
+\begin{figure}[h]
+\begin{center}
+\begin{tikzpicture}[scale=.7]
+  \begin{scope}
+    \draw (0, 0) grid (9, 9);
+    \draw[very thick, scale=3] (0, 0) grid (3, 3);
+
+    \setcounter{row}{1}
+    \foreach \i in {1,...,9}
+    {
+    \setrow {$c_{\i,1}$ }{$c_{\i,2}$}{$c_{\i,3}$}  {$c_{\i,4}$}{$c_{\i,5}$}{$c_{\i,6}$}  {$c_{\i,7}$}{$c_{\i,8}$}{$c_{\i,9}$}
+    }
+  \end{scope}
+\end{tikzpicture}
+\end{center}
+\caption{A $9\times 9$ Sudoku puzzle.}
+\end{figure}
+Thus, the Sudoku problem can be defined as:
+$$\text{sudoku}\left( \{ c_{i,j} \}_{i, j \in \{ 1,...,9\}} \right) =$$ 
+\begin{equation}\label{eq1}
+\bigwedge^9_{i=1} \text{ isvalid}\left(  c_{i,1},  c_{i,2},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9}  \right) 
+\end{equation}
+\begin{equation}\label{eq2}
+\wedge \bigwedge^9_{j=1} \text{ isvalid}\left( \right) 
+\end{equation}
+\begin{equation}\label{eq3}
+\wedge \bigwedge^9_{i,j \in \{ 1, 4, 7\}} \text{ isvalid}\left( c_{i,1},  c_{i,(j+1)},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9} \right).
+\end{equation}
+\section{Problem definition} \label{problem}
+\section{Problem definition} \label{problem}
 %%%%%

### 2021-05-23 report/paper.tex
Add problem definition, report updat

@@
 \section{Introduction} \label{introduction}
+Even though Sudoku is a Japanese word, literally meaning "digit-single", according to some sources it was actually invented in Switzerland by the already famous mathematician Euler from Basel.
+However, it is generally acknowledged that the origin of this game is European (around XIX century) and that the modern and common version has been realized in the United States in the second half of the XX century named as "Number Place", from which it spread across the globe, even in Japan, where it took the famous name it is known with.\\ \\
+The aim of this homework is to make an Encoder and Decoder for a SAT-solver in order to decide whether a problem is satisfiable or unsatisfiable. 
+\newpage
 \section{Problem definition} \label{problem}
@@
 \end{itemize}
-Often Sudoku puzzles are initialized with some values placed.
+Often Sudoku puzzles are initialized with some values placed, leaving blank spaces to be filled by the player with values respecting the rules above.
+Visually, Sudokus appear as a square grid, an example can be the following:
 
@@
 \end{figure}
-Thus, the Sudoku problem can be defined as:
-$$\text{sudoku}\left( \{ c_{i,j} \}_{i, j \in \{ 1,...,9\}} \right) =$$ 
-\begin{equation}\label{eq1}
-\bigwedge^9_{i=1} \text{ isvalid}\left(  c_{i,1},  c_{i,2},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9}  \right) 
-\end{equation}
-\begin{equation}\label{eq2}
-\wedge \bigwedge^9_{j=1} \text{ isvalid}\left( \right) 
-\end{equation}
-\begin{equation}\label{eq3}
-\wedge \bigwedge^9_{i,j \in \{ 1, 4, 7\}} \text{ isvalid}\left( c_{i,1},  c_{i,(j+1)},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9} \right).
-\end{equation}
+\noindent
+Assuming that every $c_{i,j}$ in the Sudoku grid in Figure 1 takes just a value in interval \{1,2,...,9\}, such solution can be considered valid if and only if the following conditions are satisfied: 
+\begin{enumerate}[label={(\arabic*)}]
+\item In every row $i$ (i.e., $c_{i,1}$, $c_{i,2}$, ..., $c_{i,9}$) there is at least one element that takes one value in \{1,2,...,9\}, but all values from 1 to 9 must be taken.
+Since in a row there are only 9 elements, it means that there will be a one-to-one mapping from these row elements to the \{1,2,...,9\} set values, thus each row will be made of nine different numbers.
+\item In every column $j$ (i.e., $c_{1,j}$, $c_{2,j}$, ..., $c_{9,j}$) there is at least one element that takes one value in \{1,2,...,9\}, but all values from 1 to 9 must be taken.
+Since in a column there are only 9 elements, it means that there will be a one-to-one mapping from these column elements to the \{1,2,...,9\} set values, thus each column will be made of nine different numbers.
+\item In every box, no number may appear more than once.
+Since a box is made up by 9 elements, it means that each box will contain all the values within the \{1,2,...,9\} set.
+What requires a bit more of reasoning here is finding a generalizing way to define all the elements in a box for every box.
+While for rows and columns it suffices to use a matrix-like notation, here some background reasoning is needed.
+It can be noticed that, in every box, the following pattern can be observed:
+\begin{figure}[h]
+\begin{center}
+\begin{tikzpicture}[scale=.7]
+  	\begin{scope}
+    	\draw (0, 0) grid (3, 3);
+   	\draw[very thick, scale=3] (0, 0) grid (1, 1);
+
+   	\node[anchor=center] at (0.5, 3.5) {\scriptsize{$j$}};
+   	\node[anchor=center] at (1.5, 3.5) {\scriptsize{$j+1$}};
+   	\node[anchor=center] at (2.5, 3.5) {\scriptsize{$j+2$}};
+
+   	\node[anchor=center] at (-0.5, 2.5) {\scriptsize{$i$}};
+   	\node[anchor=center] at (-0.5, 1.5) {\scriptsize{$i+1$}};
+   	\node[anchor=center] at (-0.5, 0.5) {\scriptsize{$i+2$}};
+
+   	\node[anchor=center] at (0.5, 2.5) {$c_1$};
+   	\node[anchor=center] at (1.5, 2.5) {$c_2$};
+   	\node[anchor=center] at (2.5, 2.5) {$c_3$};
+
+   	\node[anchor=center] at (0.5, 1.5) {$c_4$};
+   	\node[anchor=center] at (1.5, 1.5) {$c_5$};
+   	\node[anchor=center] at (2.5, 1.5) {$c_6$};
+
+   	\node[anchor=center] at (0.5, 0.5) {$c_7$};
+   	\node[anchor=center] at (1.5, 0.5) {$c_8$};
+   	\node[anchor=center] at (2.5, 0.5) {$c_9$};
+
+  	\end{scope}
+\end{tikzpicture}
+\end{center}
+\caption{A $3\times 3$ Sudoku generalized box.}
+\end{figure}
+
+\noindent
+That means that every box has its top-leftmost cell at a particular coordinate ($i,\,j$) and its other corner cells are clockwise respectively at coordinates: ($i,\,j+2$),  ($i+2,\,j+2$) and ($i+2,\,j$).
+By this, it can be said that by individuating the top-leftmost cell of a box it is possible to individuate the box itself.
+Hence, for $i=1$ and $j=1$, we have corner cells ($1,\,1$), ($1,\,3$),  ($3,\,3$) and ($3,\,1$), that is the top-leftmost box in the Sudoku grid.
+From this approach, the boxes making up the grid can be defined as:
+\begin{figure}[h]
+\begin{center}
+\begin{tikzpicture}[scale=1]
+  	\begin{scope}
+    	\draw (0, 0) grid (3, 3);
+   	\draw[very thick, scale=3] (0, 0) grid (1, 1);
+
+   	\node[anchor=center] at (0.5, 2.5) {$(1,1)$};
+   	\node[anchor=center] at (1.5, 2.5) {$(1,4)$};
+   	\node[anchor=center] at (2.5, 2.5) {$(1,7)$};
+
+   	\node[anchor=center] at (0.5, 1.5) {$(4,1)$};
+   	\node[anchor=center] at (1.5, 1.5) {$(4,4)$};
+   	\node[anchor=center] at (2.5, 1.5) {$(4,7)$};
+
+   	\node[anchor=center] at (0.5, 0.5) {$(7,1)$};
+   	\node[anchor=center] at (1.5, 0.5) {$(7,4)$};
+   	\node[anchor=center] at (2.5, 0.5) {$(7,7)$};
+
+  	\end{scope}
+\end{tikzpicture}
+\end{center}
+\caption{A $9\times 9$ Sudoku grid. Each square corresponds to a $3\times 3$ box. Inside of the square, the ($i,\,j$) coordinate of the top-leftmost cell of that box is written.}
+\end{figure}
+
+\noindent
+It can be seen that the coordinates are nothing but all the possible permutations of three values: 1, 4 and 7. From this, we can define a set $B=\{1,\,4,\,7\}$ containing these values.\\
+\end{enumerate}
+\noindent
+Let \textbf{isvalid} define whether a row, column or box is valid. That means that the nine elements along a row, column or within a box must be different and equal to a number in set $\{1,2,...,9\}$.
+\begin{align}
+\text{isvalid}\left(  c_1,  c_2,  c_3,  c_4,  c_5,  c_6,  c_7,  c_8,  c_9  \right) = \bigwedge^9_{i=1}  \bigvee^9_{i=1} \,\,\, c_i
+\end{align}
+
+\noindent
+Therefore, the Sudoku problem can be defined as:
+\setcounter{equation}{0}
+\begin{align}
+\text{sudoku}\left( \{ c_{i,j} \}_{i, j \in \{ 1,...,9\}} \right) = & \nonumber\\
+&\bigwedge^9_{i=1} \text{ isvalid}\left(  c_{i,1},  c_{i,2},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9}  \right)\\
+\wedge &\bigwedge^9_{j=1} \text{ isvalid}\left( c_{1,j},  c_{2,j},  c_{3,j},  c_{4,j},  c_{5,j},  c_{6,j},  c_{7,j},  c_{8,j},  c_{9,j} \right) \\
+\wedge &\bigwedge^9_{i,j \in B} \text{ isvalid}\left( c_{i,1},  c_{i,(j+1)},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9} \right).
+\end{align}
 \section{Problem definition} \label{problem}

### 2021-05-24 solver/app.py
The wall of berlin has fallen

@@ -3,6 +3,12 @@ from z3 import And, Int, Or, Solver, sat
 
 
 def is_valid(*data: str):
+    """
+    Create SAT condition that ensures
+    all the entries in the data array are
+    different from one another (supports
+    mixed pre-filled and unknown variables)
+    """
     x = []
     for i in range(0, 9):
         if data[i].isnumeric():
@@ -20,28 +26,32 @@ def is_valid(*data: str):
     return And(*and_args)
 
 
-def sudoku_solver(sudoku):
+def build_sudoku_solver(board):
+    """
+    Return a solver for a
+    given sudoku board
+    """
     rows_check = []
     for i in range(0, 9):
-        rows_check.append(is_valid(sudoku[i][0], sudoku[i][1], sudoku[i][2],
-                                   sudoku[i][3], sudoku[i][4], sudoku[i][5],
-                                   sudoku[i][6], sudoku[i][7], sudoku[i][8]))
+        rows_check.append(is_valid(board[i][0], board[i][1], board[i][2],
+                                   board[i][3], board[i][4], board[i][5],
+                                   board[i][6], board[i][7], board[i][8]))
     columns_check = []
     for i in range(0, 9):
-        columns_check.append(is_valid(sudoku[0][i], sudoku[1][i], sudoku[2][i],
-                                      sudoku[3][i], sudoku[4][i], sudoku[5][i],
-                                      sudoku[6][i], sudoku[7][i],
-                                      sudoku[8][i]))
+        columns_check.append(is_valid(board[0][i], board[1][i], board[2][i],
+                                      board[3][i], board[4][i], board[5][i],
+                                      board[6][i], board[7][i],
+                                      board[8][i]))
     squares_check = []
     for i in range(0, 9, 3):
         for j in range(0, 9, 3):
-            squares_check.append(is_valid(sudoku[i][j], sudoku[i][j + 1],
-                                          sudoku[i][j + 2], sudoku[i + 1][j],
-                                          sudoku[i + 1][j + 1],
-                                          sudoku[i + 1][j + 2],
-                                          sudoku[i + 2][j],
-                                          sudoku[i + 2][j + 1],
-                                          sudoku[i + 2][j + 2]))
+            squares_check.append(is_valid(board[i][j], board[i][j + 1],
+                                          board[i][j + 2], board[i + 1][j],
+                                          board[i + 1][j + 1],
+                                          board[i + 1][j + 2],
[…]

### 2021-05-24 solver/socrates.py

@@ -1,26 +1,0 @@
-from z3 import *
-
-Object = DeclareSort('Object')
-Human = Function('Human', Object, BoolSort())
-Mortal = Function('Mortal', Object, BoolSort())
-
-# A well known philosopher 
-socrates = Const('socrates', Object)
-
-# Free variables used in forall must be declared Const in python
-x = Const('x', Object)
-
-axioms = [ForAll([x], Implies(Human(x), Mortal(x))), 
-          Human(socrates)]
-
-s = Solver()
-s.add(axioms)
-
-# prints sat so axioms are coherent
-print(s.check())
-
-# classical refutation
-s.add(Not(Mortal(socrates)))
-
-# prints unsat so socrates is Mortal
-print(s.check()) 

### 2021-05-24 report/paper.tex
Report style update
Add important things

@@
 \usepackage{enumitem}
+\usepackage{listings}
+\usepackage[ruled]{algorithm}
+\usepackage[noend]{algpseudocode}
+
+\lstdefinestyle{bpstyle}{
+  literate={*}{*}1
+}
+
+\lstset{
+  style=bpstyle
+}
 
@@
 \section{Introduction} \label{introduction}
-Even though Sudoku is a Japanese word, literally meaning "digit-single", according to some sources it was actually invented in Switzerland by the already famous mathematician Euler from Basel.
+Even though Sudoku is a Japanese word, literally meaning "single digit", according to some sources it was actually invented in Switzerland by the already famous mathematician Euler from Basel.
 However, it is generally acknowledged that the origin of this game is European (around XIX century) and that the modern and common version has been realized in the United States in the second half of the XX century named as "Number Place", from which it spread across the globe, even in Japan, where it took the famous name it is known with.\\ \\
@@
 \item In every cell there is exactly one number $c \in \{1, 2, ..., 9 \}$
-\item No number may appear more than once in the same row or column.
+\item No number may appear more than once in the same row.
+\item No number may appear more than once in the same column.
 \item No number may appear more than once in the same box.
@@
 \end{enumerate}
+\newpage
+\section{Encoder and Decoder}\label{encdec}
 \noindent
-Let \textbf{isvalid} define whether a row, column or box is valid. That means that the nine elements along a row, column or within a box must be different and equal to a number in set $\{1,2,...,9\}$.
-\begin{align}
-\text{isvalid}\left(  c_1,  c_2,  c_3,  c_4,  c_5,  c_6,  c_7,  c_8,  c_9  \right) = \bigwedge^9_{i=1}  \bigvee^9_{i=1} \,\,\, c_i
-\end{align}
+Let \textbf{is\_valid} define whether a row, column or box is valid.
+That means that the nine elements along a row, column or within a box must be different and equal to a digit in set $\{1,2,...,9\}$, thus meaning that a row, column or box is valid if it contains all digits.
+Let's consider $c_1,\,c_2,\,...,\,c_9$, which may be the elements within a row, column or box.
+Then we have:
+\begin{align*}
+\text{is\_valid}\left(  c_1,  c_2,  c_3,  c_4,  c_5,  c_6,  c_7,  c_8,  c_9  \right) = & \\
+& \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
+& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
+& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
+& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
+& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
+& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
+& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
+& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
+& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
+= & \bigwedge^9_{i=1}  \bigvee^9_{i=1} \,\,\, c_i
+\end{align*}
 
 \noindent
-Therefore, the Sudoku problem can be defined as:
+Now, since we have given a definition of valid, we can say that a Sudoku is valid if and only if: 
+\begin{enumerate}[label={(\arabic*)}]
+\item Every row is valid.
+\item Every column is valid.
+\item Every box is valid.
+\end{enumerate}
+
+\noindent
+Therefore, the Sudoku problem for a 9x9 grid can be defined as:
 \setcounter{equation}{0}
@@
 \text{sudoku}\left( \{ c_{i,j} \}_{i, j \in \{ 1,...,9\}} \right) = & \nonumber\\
-&\bigwedge^9_{i=1} \text{ isvalid}\left(  c_{i,1},  c_{i,2},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9}  \right)\\
-\wedge &\bigwedge^9_{j=1} \text{ isvalid}\left( c_{1,j},  c_{2,j},  c_{3,j},  c_{4,j},  c_{5,j},  c_{6,j},  c_{7,j},  c_{8,j},  c_{9,j} \right) \\
-\wedge &\bigwedge^9_{i,j \in B} \text{ isvalid}\left( c_{i,1},  c_{i,(j+1)},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9} \right).
+&\bigwedge^9_{i=1} \text{ is\_valid}\left(  c_{i,1},  c_{i,2},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9}  \right)\\
+\wedge &\bigwedge^9_{j=1} \text{ is\_valid}\left( c_{1,j},  c_{2,j},  c_{3,j},  c_{4,j},  c_{5,j},  c_{6,j},  c_{7,j},  c_{8,j},  c_{9,j} \right) \\
+\wedge &\bigwedge^9_{i,j \in \{1,\,4,\,7\}} \text{ is\_valid}\left( c_{i,1},  c_{i,(j+1)},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9} \right).
 \end{align}
-\section{Problem definition} \label{problem}
-\section{Problem definition} \label{problem}
+\newpage
+\section{Development}\label{development}
+We have developed the project in Python3 to provide both a web interface and the solver logic using z3.\\
+For what concerns the z3 SAT solver, we have used the official \texttt{z3–solver} Python3 library to write the z3 logic instead of writing a z3 \texttt{.dimacs} file for every execution and invoking the z3 solver using a system shell.
+This allows for better performance and easier maintenance of the code.
+The \texttt{z3–solver} Python3 library allows to specify constraints programmatically and is able to provide both information about the satisfiability information and data structure models that represent a valid solution for the problem.\\ \\
+First of all, we want to translate the following definition into Python3 \texttt{z3–solver} code:
+\begin{align*}
+\text{sudoku}\left( \{ c_{i,j} \}_{i, j \in \{ 1,...,9\}} \right) = & \nonumber\\
+       & \bigwedge^9_{i=1} \text{ is\_valid}\left(  c_{i,1},  c_{i,2},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9}  \right)\\
+\wedge &\bigwedge^9_{j=1} \text{ is\_valid}\left( c_{1,j},  c_{2,j},  c_{3,j},  c_{4,j},  c_{5,j},  c_{6,j},  c_{7,j},  c_{8,j},  c_{9,j} \right) \\
+\wedge &\bigwedge^9_{i,j \in B} \text{ is\_valid}\left( c_{i,1},  c_{i,(j+1)},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9} \right).
+\end{align*}
+
+\noindent
+Here we build three list of conditions that should all be satisfied: the first condition guarantees that each row does not contain any repeated element within itself.
+The second condition guarantees that each column does not contain any repeated element while the third one, as described previously
+in section~\ref{problem}, guarantees the uniqueness of each element in the 9 boxes of our Sudoku board.
+Now, it follows the pseudo-code algorithm that we designed and was used as a basis for the \texttt{z3–solver} API in the Python3 implementation.
+\begin{algorithm}[H]
+\hspace*{\algorithmicindent} \textbf{Input}: $c$, Sudoku board. \\
+\hspace*{\algorithmicindent} \textbf{Output}: $s$, solver. 
+\begin{algorithmic}[1]\label{algorithm:}
+\Function {BuildSolver}{$c$}
+\State {\textsc{RowsCheck} $\gets$ []} \Comment{Create array for rows check}
+\For {$i$ in range $[1,9]$}\Comment{Go across nine elements}
+\State {\textsc{RowsCheck}.\textsc{push}$\left(\textsc{isValid}(c_{1,i},\, c_{2,i},\, c_{3,i},\, c_{4,i},\, c_{5,i},\, c_{6,i},\, c_{7,i},\, c_{8,i},\, c_{9,i}) \right)$}
+\EndFor
+\State {\textsc{ColumnsCheck} $\gets$ []} \Comment{Create array for columns check}
+\For {$i$ in range $[1,9]$}\Comment{Go across nine elements}
+\State {\textsc{ColumnsCheck}.\textsc{push}$\left(\textsc{isValid}(c_{i,1},\, c_{i,2},\, c_{i,3},\, c_{i,4},\, c_{i,5},\, c_{i,6},\, c_{i,7},\, c_{i,8},\, c_{i,9}) \right)$}
+\EndFor
+\State {\textsc{BoxesCheck} $\gets$ []} \Comment{Create array for boxes check}
+\For {$i$ in range $[1,9]$ step = 3}
+\For {$j$ in range $[1,9]$ step = 3}
+\State {\textsc{BoxesCheck}.\textsc{push}$\left(\textsc{isValid}(c_{i,j},\, c_{i,j+1},\, c_{i,j+2},\, c_{1+1,j},\, c_{i+1,j+1},\, c_{i+1,j+2},\, c_{i+2,j},\, c_{i+2,j+1},\, c_{i+2,j+2}) \right)$}
+\EndFor
+\EndFor
+\State{\textsc{s} $\gets$ \textsc{Z3Solver()}}
+\State{\textsc{s}.\textsc{add}$\left( \textsc{And} \left( \textsc{And} \left( \textsc{RowsCheck}\right),\, \textsc{And} \left( \textsc{ColumnsCheck}\right),\,\textsc{And} \left( \textsc{BoxesCheck}\right) \right) \right)$}\\
+\Return {$s$}
+\EndFunction
+\end{algorithmic}
+\caption {Build Solver Algorithm}
+\end{algorithm}
+
+\newpage
+\noindent
+Now we want to write the $ is\_valid $ formulation as Python3 \texttt{z3–solver}
+code:
+
+\begin{align*}
+\text{is\_valid}\left(  c_1,  c_2,  c_3,  c_4,  c_5,  c_6,  c_7,  c_8,  c_9  \right) = \bigwedge^9_{i=1}  \bigvee^9_{i=1} \,\,\, c_i
+\end{align*}
+
+\begin{algorithm}[H]
+\hspace*{\algorithmicindent} \textbf{Input}: $c$, Sudoku board. \\
+\hspace*{\algorithmicindent} \textbf{Output}: $s$, solver. 
+\begin{algorithmic}[1]\label{algorithm:}
+\Function {isValid}{$c$}
+\State {\textsc{x} $\gets$ []} \Comment{Create array}
+\For {$i$ in range [1,9]}
+\If {$c_i$.\textsc{isNumeric()}}
+\State{\textsc{x.push}($c_i$)}
+\Else
+\State{\textsc{x.push}($c_i$)}
+\EndIf
+\EndFor
+\For {$d$ in range [1,9]}
+\State {\textsc{orArgs $\gets$ []}}
+\For {$d$ in range [1,9]}
+\State {\textsc{orArgs.push($x_i == d$)}}
+\EndFor
+\State {\textsc{andArgs.push(Or(orArgs))}}
+\EndFor\\
+\Return {$\textsc{And(andArgs)}$}
+\EndFunction
+\end{algorithmic}
+\caption {Build Solver Algorithm}
+\end{algorithm}
+
+\newpage
+\section{Instructions for running}
+
+In order to run the Sudoku SAT solver, you first need to install Python3 (available at \url{https://www.python.org/downloads/}).
+Then you need to open the terminal and browse to the project directory using the \texttt{cd}
+command.
+Once you have reached the destination folder which is the project directory, run the following command to install the dependencies using the Python3 \texttt{pip} command (that literally stands for ``pip install package'', according to its \href{https://en.wikipedia.org/wiki/Pip\_(package\_manager)}{Wikipedia page}):
+
+\begin{center}
+\texttt{pip install -r requirements.txt}
+\end{center}
+
+\noindent
+Finally, now that you have installed the dependencies, you can now start the Flask server.
+Flask is a web server framework for Python3, see the \href{https://flask.palletsprojects.com/}{official documentation}.\\ \\
+Run this command to start the Sudoku SAT Solver:
+
+\begin{center}
+\texttt{FLASK\_APP=solver/app flask run}
+\end{center}
+
+\noindent
+And finally open the \url{http://localhost:5000/} website to visit the local
+instance of the Sudoku SAT Solver project.
+
+\section{Trials} \label{trials}
 %%%%%

### 2021-05-25 report/paper.tex
Add very important pictures and show that we can solve the world's hardest sudoku
Signed-off-by: Joey <bevilacquajoey@gmail.com>

@@
 
-\section{Trials} \label{trials}
+\section{Trials}\label{trials}
+
+\begin{figure}
+\caption{The web app UI}
+\centering
+\includegraphics[width=0.8\textwidth]{pics/app_ui.png}
+\end{figure}
+
+\begin{figure}
+\caption{Fill the sudoku board}
+\centering
+\includegraphics[width=0.8\textwidth]{pics/fill_board.png}
+\end{figure}
+
+\begin{figure}
+\caption{Press the \texttt{Solve} button to find the soluscion}
+\centering
+\includegraphics[width=0.8\textwidth]{pics/solved.png}
+\end{figure}
+
+\begin{figure}
+\caption{Errors if a number appears multiple times in the same row}
+\centering
+\includegraphics[width=0.8\textwidth]{pics/row_check.png}
+\end{figure}
+
+\begin{figure}
+\caption{Errors if a number appears multiple times in the same column}
+\centering
+\includegraphics[width=0.8\textwidth]{pics/column_check.png}
+\end{figure}
+
+\begin{figure}
+\caption{Errors if a number appears multiple times in the same box}
+\centering
+\includegraphics[width=0.8\textwidth]{pics/box_check.png}
+\end{figure}
+
+\begin{figure}
+\caption{Errors if a number is not in $ \left[ 1, 9 \right] $}
+\centering
+\includegraphics[width=0.8\textwidth]{pics/column_check.png}
+\end{figure}
+
+\begin{figure}
+\caption{Errors if a number is not in $ \left[ 1, 9 \right] $}
+\centering
+\includegraphics[width=0.8\textwidth]{pics/input_check.png}
+\end{figure}
+
+\begin{figure}
+\caption{Solve \href{https://www.mirror.co.uk/news/weird-news/worlds-hardest-sudoku-can-you-242294}{the world's hardest sudoku} in seconds}
+\centering
+\includegraphics[width=0.8\textwidth]{pics/hardest_sudoku.png}
+\end{figure}
+
 %%%%%

### 2021-05-25 report/paper.tex
(Literally) fix the figures
Signed-off-by: Joey <bevilacquajoey@gmail.com>
malusa

@@
 \usepackage[noend]{algpseudocode}
+\usepackage[section]{placeins}
 
@@
 \begin{itemize}%[label={(\arabic*)}]
-\item In every cell there is exactly one number $c \in \{1, 2, ..., 9 \}$
+\item In every cell there is exactly one number $c \in \{1,2,3,4,5,6,7,8,5,6,7,8, 9 \}$
 \item No number may appear more than once in the same row.
@@
 \noindent
-Assuming that every $c_{i,j}$ in the Sudoku grid in Figure 1 takes just a value in interval \{1,2,...,9\}, such solution can be considered valid if and only if the following conditions are satisfied: 
+Assuming that every $c_{i,j}$ in the Sudoku grid in Figure 1 takes just a value in interval \{1,2,3,4,5,6,7,8,9\}, such solution can be considered valid if and only if the following conditions are satisfied:
 \begin{enumerate}[label={(\arabic*)}]
-\item In every row $i$ (i.e., $c_{i,1}$, $c_{i,2}$, ..., $c_{i,9}$) there is at least one element that takes one value in \{1,2,...,9\}, but all values from 1 to 9 must be taken.
-Since in a row there are only 9 elements, it means that there will be a one-to-one mapping from these row elements to the \{1,2,...,9\} set values, thus each row will be made of nine different numbers.
-\item In every column $j$ (i.e., $c_{1,j}$, $c_{2,j}$, ..., $c_{9,j}$) there is at least one element that takes one value in \{1,2,...,9\}, but all values from 1 to 9 must be taken.
-Since in a column there are only 9 elements, it means that there will be a one-to-one mapping from these column elements to the \{1,2,...,9\} set values, thus each column will be made of nine different numbers.
+\item In every row $i$ (i.e., $c_{i,1}$, $c_{i,2}$, ..., $c_{i,9}$) there is at least one element that takes one value in \{1,2,3,4,5,6,7,8,9\}, but all values from 1 to 9 must be taken.
+Since in a row there are only 9 elements, it means that there will be a one-to-one mapping from these row elements to the \{1,2,3,4,5,6,7,8,9\} set values, thus each row will be made of nine different numbers.
+\item In every column $j$ (i.e., $c_{1,j}$, $c_{2,j}$, ..., $c_{9,j}$) there is at least one element that takes one value in \{1,2,3,4,5,6,7,8,9\}, but all values from 1 to 9 must be taken.
+Since in a column there are only 9 elements, it means that there will be a one-to-one mapping from these column elements to the \{1,2,3,4,5,6,7,8,9\} set values, thus each column will be made of nine different numbers.
 \item In every box, no number may appear more than once.
-Since a box is made up by 9 elements, it means that each box will contain all the values within the \{1,2,...,9\} set.
+Since a box is made up by 9 elements, it means that each box will contain all the values within the \{1,2,3,4,5,6,7,8,9\} set.
 What requires a bit more of reasoning here is finding a generalizing way to define all the elements in a box for every box.
@@
 Let \textbf{is\_valid} define whether a row, column or box is valid.
-That means that the nine elements along a row, column or within a box must be different and equal to a digit in set $\{1,2,...,9\}$, thus meaning that a row, column or box is valid if it contains all digits.
+That means that the nine elements along a row, column or within a box must be different and equal to a digit in set $\{1,2,3,4,5,6,7,8,9\}$, thus meaning that a row, column or box is valid if it contains all digits.
 Let's consider $c_1,\,c_2,\,...,\,c_9$, which may be the elements within a row, column or box.
@@
 \begin{align*}
-\text{is\_valid}\left(  c_1,  c_2,  c_3,  c_4,  c_5,  c_6,  c_7,  c_8,  c_9  \right) = & \\
-& \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
-& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
-& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
-& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
-& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
-& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
-& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
-& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
-& \wedge \left(  c_1 \vee  c_2 \vee c_3 \vee c_4 \vee  c_5 \vee  c_6 \vee  c_7 \vee  c_8 \vee  c_9  \right)\\
-= & \bigwedge^9_{i=1}  \bigvee^9_{i=1} \,\,\, c_i
+\text{is\_valid}\left(  c_1,  c_2,  c_3,  c_4,  c_5,  c_6,  c_7,  c_8,  c_9
+	\right) \Leftrightarrow & \\
+& \left(  c_1  = 1 \vee  c_2  = 1 \vee c_3  = 1 \vee c_4  = 1 \vee  c_5  = 1
+	\vee  c_6  = 1 \vee  c_7  = 1 \vee  c_8  = 1 \vee  c_9 = 1 \right)\\
+& \wedge \left(  c_1  = 2 \vee  c_2  = 2 \vee c_3  = 2 \vee c_4  = 2 \vee  c_5
+	= 2 \vee  c_6  = 2 \vee  c_7  = 2 \vee  c_8  = 2 \vee  c_9 = 2  \right)\\
+& \wedge \left(  c_1  = 3 \vee  c_2  = 3 \vee c_3  = 3 \vee c_4  = 3 \vee  c_5
+	= 3 \vee  c_6  = 3 \vee  c_7  = 3 \vee  c_8  = 3 \vee  c_9 = 3 \right)\\
+& \wedge \left(  c_1  = 4 \vee  c_2  = 4 \vee c_3  = 4 \vee c_4  = 4 \vee  c_5
+	= 4 \vee  c_6  = 4 \vee  c_7  = 4 \vee  c_8  = 4 \vee  c_9 = 4 \right)\\
+& \wedge \left(  c_1  = 5 \vee  c_2  = 5 \vee c_3  = 5 \vee c_4  = 5 \vee  c_5
+	= 5 \vee  c_6  = 5 \vee  c_7  = 5 \vee  c_8  = 5 \vee  c_9 = 5 \right)\\
+& \wedge \left(  c_1  = 6 \vee  c_2  = 6 \vee c_3  = 6 \vee c_4  = 6 \vee  c_5
+	= 6 \vee  c_6  = 6 \vee  c_7  = 6 \vee  c_8  = 6 \vee  c_9 = 6 \right)\\
+& \wedge \left(  c_1  = 7 \vee  c_2  = 7 \vee c_3  = 7 \vee c_4  = 7 \vee  c_5
+	= 7 \vee  c_6  = 7 \vee  c_7  = 7 \vee  c_8  = 7 \vee  c_9 = 7 \right)\\
+& \wedge \left(  c_1  = 8 \vee  c_2  = 8 \vee c_3  = 8 \vee c_4  =  8 \vee  c_5
+	= 8 \vee  c_6  = 8 \vee  c_7  = 8 \vee  c_8  = 8 \vee  c_9 = 8 \right)\\
+& \wedge \left(  c_1  = 9 \vee  c_2  = 9 \vee c_3  = 9 \vee c_4  = 9 \vee  c_5
+	= 9 \vee  c_6  = 9 \vee  c_7  = 9 \vee  c_8  = 9 \vee  c_9 = 9  \right)\\
+\Leftrightarrow & \bigwedge^9_{n=1}  \bigvee^9_{i=1} \,\,\, c_i = n
 \end{align*}
@@
 \noindent
-Therefore, the Sudoku problem for a 9x9 grid can be defined as:
+Therefore, the Sudoku problem for a 9x9 grid can be defined as:\\
+
 \setcounter{equation}{0}
-\begin{align}
-\text{sudoku}\left( \{ c_{i,j} \}_{i, j \in \{ 1,...,9\}} \right) = & \nonumber\\
+$\text{sudoku}\left( \{ c_{i,j} \}_{i, j \in \{ 1,2,3,4,5,6,7,8,9\}} \right) = $\begin{align}
 &\bigwedge^9_{i=1} \text{ is\_valid}\left(  c_{i,1},  c_{i,2},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9}  \right)\\
 \wedge &\bigwedge^9_{j=1} \text{ is\_valid}\left( c_{1,j},  c_{2,j},  c_{3,j},  c_{4,j},  c_{5,j},  c_{6,j},  c_{7,j},  c_{8,j},  c_{9,j} \right) \\
-\wedge &\bigwedge^9_{i,j \in \{1,\,4,\,7\}} \text{ is\_valid}\left( c_{i,1},  c_{i,(j+1)},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9} \right).
+	\wedge &\bigwedge^9_{i,j \in \{1,\,4,\,7\}} \text{ is\_valid}\left( c_{i,1},
+	c_{i,(j+1)},  c_{i,(j+2)}, c_{(i+1),j},  c_{(i+1),(j+1)},
+	c_{(i+1), (j+2)}, c_{(i+2),j},  c_{(i+2),(j+1)},  c_{(i+2),(j+2)} \right).
 \end{align}
@@
 \begin{align*}
-\text{is\_valid}\left(  c_1,  c_2,  c_3,  c_4,  c_5,  c_6,  c_7,  c_8,  c_9  \right) = \bigwedge^9_{i=1}  \bigvee^9_{i=1} \,\,\, c_i
+\text{is\_valid}\left(  c_1,  c_2,  c_3,  c_4,  c_5,  c_6,  c_7,  c_8,  c_9
+	\right) = \bigwedge^9_{i=1}  \bigvee^9_{i=1} \,\,\, c_i
 \end{align*}
@@
 
-In order to run the Sudoku SAT solver, you first need to install Python3 (available at \url{https://www.python.org/downloads/}).
-Then you need to open the terminal and browse to the project directory using the \texttt{cd}
-command.
-Once you have reached the destination folder which is the project directory, run the following command to install the dependencies using the Python3 \texttt{pip} command (that literally stands for ``pip install package'', according to its \href{https://en.wikipedia.org/wiki/Pip\_(package\_manager)}{Wikipedia page}):
+In order to run the Sudoku SAT solver, you first need to install Python3
+(available at \url{https://www.python.org/downloads/}). Then you need to open
+the terminal and browse to the project directory using the \texttt{cd} command.
+Once you have reached the destination folder which is the project directory, run
+the following command to install the dependencies using the Python3 \texttt{pip}
+command (that literally stands for ``pip install package'', according to its
+\href{https://en.wikipedia.org/wiki/Pip\_(package\_manager)}{Wikipedia page}):
 
@@
 
-\noindent
-Finally, now that you have installed the dependencies, you can now start the Flask server.
-Flask is a web server framework for Python3, see the \href{https://flask.palletsprojects.com/}{official documentation}.\\ \\
-Run this command to start the Sudoku SAT Solver:
+\noindent Finally, now that you have installed the dependencies, you can now
+start the Flask server. Flask is a web server framework for Python3, see the
+\href{https://flask.palletsprojects.com/}{official documentation}.\\ \\ Run this
+command to start the Sudoku SAT Solver:
 
@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.8\textwidth]{pics/app_ui.png}
+\caption{The web app UI}
+\end{figure}
+
+\newpage
+
 \section{Trials}\label{trials}
 
-\begin{figure}
-\caption{The web app UI}
+\begin{figure}[ht]
 \centering
-\includegraphics[width=0.8\textwidth]{pics/app_ui.png}
+\includegraphics[width=0.8\textwidth]{pics/fill_board.png}
+\caption{Fill the sudoku board}
 \end{figure}
 
-\begin{figure}
-\caption{Fill the sudoku board}
+\begin{figure}[ht]
 \centering
-\includegraphics[width=0.8\textwidth]{pics/fill_board.png}
+\includegraphics[width=0.8\textwidth]{pics/solved.png}
+\caption{Press the \texttt{Solve} button to find the soluscion}
 \end{figure}
 
-\begin{figure}
-\caption{Press the \texttt{Solve} button to find the soluscion}
+\begin{figure}[ht]
 \centering
-\includegraphics[width=0.8\textwidth]{pics/solved.png}
+\includegraphics[width=0.8\textwidth]{pics/row_check.png}
+\caption{Errors if a number appears multiple times in the same row}
 \end{figure}
 
-\begin{figure}
-\caption{Errors if a number appears multiple times in the same row}
+\begin{figure}[ht]
 \centering
-\includegraphics[width=0.8\textwidth]{pics/row_check.png}
+\includegraphics[width=0.8\textwidth]{pics/column_check.png}
+\caption{Errors if a number appears multiple times in the same column}
 \end{figure}
 
-\begin{figure}
-\caption{Errors if a number appears multiple times in the same column}
+\begin{figure}[ht]
 \centering
-\includegraphics[width=0.8\textwidth]{pics/column_check.png}
+\includegraphics[width=0.8\textwidth]{pics/box_check.png}
+\caption{Errors if a number appears multiple times in the same box}
 \end{figure}
 
-\begin{figure}
-\caption{Errors if a number appears multiple times in the same box}
+\begin{figure}[ht]
 \centering
-\includegraphics[width=0.8\textwidth]{pics/box_check.png}
+\includegraphics[width=0.8\textwidth]{pics/input_check.png}
+\caption{Errors if a number is not in $ \left[ 1, 9 \right] $}
 \end{figure}
 
-\begin{figure}
-\caption{Errors if a number is not in $ \left[ 1, 9 \right] $}
-\centering
-\includegraphics[width=0.8\textwidth]{pics/column_check.png}
-\end{figure}
-
-\begin{figure}
-\caption{Errors if a number is not in $ \left[ 1, 9 \right] $}
-\centering
-\includegraphics[width=0.8\textwidth]{pics/input_check.png}
-\end{figure}
-
-\begin{figure}
-\caption{Solve \href{https://www.mirror.co.uk/news/weird-news/worlds-hardest-sudoku-can-you-242294}{the world's hardest sudoku} in seconds}
+\begin{figure}[ht]
 \centering
 \includegraphics[width=0.8\textwidth]{pics/hardest_sudoku.png}
+\caption{Solve \href{https://www.mirror.co.uk/news/weird-news/worlds-hardest-sudoku-can-you-242294}{the world's hardest sudoku} in seconds}
 \end{figure}

### 2021-05-25 report/paper.tex
Igor Moreno Santos

@@
 \tableofcontents
+\listoffigures
 \newpage
@@
 The aim of this homework is to make an Encoder and Decoder for a SAT-solver in order to decide whether a problem is satisfiable or unsatisfiable.
-\newpage
-\section{Problem definition} \label{problem}
+\subsection{Problem definition} \label{problem}
 Sudokus are puzzles based upon number placement.
@@
 \begin{itemize}%[label={(\arabic*)}]
-\item In every cell there is exactly one number $c \in \{1,2,3,4,5,6,7,8,5,6,7,8, 9 \}$
+\item In every cell there is exactly one number $c \in \{1,2,3,4,5,6,7,8,9 \}$
 \item No number may appear more than once in the same row.
@@
 \end{align*}
+\hspace{0cm}\\
+\noindent
+This definition follows by the pigeon hole principle. Our aim is to check if the
+nine inputs of \texttt{is\_valid} contain each one digit in
+$\{1,2,3,4,5,6,7,8,9\}$. Each of the OR-clauses enclosed in brackets check if
+each digit is present at least one time, but since we have 9 clauses it follows
+that if even a single duplicate digit is given then at least an OR-clause will
+have false value. If we imagine each OR-clause is a ``hole'' in which we deposit
+digit, in this case the ``hole'' corresponding to the duplicate digit will be
+``fed'' twice, and since we have only 9 ``holes'' and 9 ``pigeons'', the
+duplicate will cause a ``hole'' to be void of ``pigeons'', thus making the
+entire \texttt{is\_valid} formula false and correctly rejecting the invalid
+input.
+\hspace{0cm}\\
 
@@
 in section~\ref{problem}, guarantees the uniqueness of each element in the 9 boxes of our Sudoku board.
-Now, it follows the pseudo-code algorithm that we designed and was used as a basis for the \texttt{z3–solver} API in the Python3 implementation.
+
+\subsection{The build solver algorithm}
+Below it follows the pseudo-code algorithm that we designed and was used as a basis for the \texttt{z3–solver} API in the Python3 implementation.
 \begin{algorithm}[H]
@@
 
-\newpage
+\subsection{The \texttt{is\_valid} function}
 \noindent
@@
 \end{algorithmic}
-\caption {Build Solver Algorithm}
+	\caption {\texttt{is\_valid} Algorithm}
 \end{algorithm}
@@
 
-\section{Trials}\label{trials}
+\section{Tests and examples}\label{trials}
+
+In this section we aim to test the Sudoku SAT solver with some Sudoku examples
+and edge cases in order to better show the algorithm and GUI's behaviour.
+
+Please
+find in the figure below the Sudoku examples.
 
@@
 \includegraphics[width=0.8\textwidth]{pics/input_check.png}
-\caption{Errors if a number is not in $ \left[ 1, 9 \right] $}
+\caption{Errors if a number is not in $ \left[ 1, 2,3,4,5,6,7,8, 9 \right] $}
 \end{figure}

### 2021-05-25 report/paper.tex

@@
 \abstract {
-The aim of this project is to make an Encoder and Decoder for a SAT-solver in order to decide whether a problem is satisfiable or unsatisfiable. WRITE A BETTER ABSTRACT
+Sudoku is a Swiss game often confused as japanese because of its japanese
+sounding name.
+
+Sudokus are puzzles composed by numbers that are structured in a $9 \times 9$ table.
+At the beginning of the game the Sudoku board only a few cells contain numbers,
+which are ``randomly'' positioned in
+the table.
+
+The main purpose of the game is to fill the table without repeating numbers
+in a ``box'' or in the entire row or column of the table.
+
+The aim of this
+project is to make an Encoder and Decoder for a SAT-solver in order to decide
+whether a given input board is solvable or not, and when solvable to find a
+feasible solution.
+
+A Sudoku is valid when all of its rows are valid
+(each row must not have repeating numbers), when all of its columns
+are valid (each column must not have repeating numbers),
+and when all of the $3 \times 3$ boxes in the board are valid
+(each box must not have repeating numbers).
+
+To develop the code of the project
+we used Python3, specifically using the pip3 package z3-solver (official
+library to write logics for SAT solver).
 }

### 2021-05-25 report/paper.tex
oiafpbiuaspibfsa

@@
     \setcounter{row}{1}
-    \foreach \i in {1,...,9}
+    \foreach \i in {1,2,3,4,5,6,7,8,9}
     {
@@
 \begin{enumerate}[label={(\arabic*)}]
-\item In every row $i$ (i.e., $c_{i,1}$, $c_{i,2}$, ..., $c_{i,9}$) there is at least one element that takes one value in \{1,2,3,4,5,6,7,8,9\}, but all values from 1 to 9 must be taken.
+\item In every row $i$ (i.e., $c_{i,1}$, $c_{i,2}$, $c_{i,3}$, $c_{i,4}$, $c_{i,5}$, $c_{i,6}$, $c_{i,7}$, $c_{i,8}$, $c_{i,9}$) there is at least one element that takes one value in \{1,2,3,4,5,6,7,8,9\}, but all values from 1 to 9 must be taken.
 Since in a row there are only 9 elements, it means that there will be a one-to-one mapping from these row elements to the \{1,2,3,4,5,6,7,8,9\} set values, thus each row will be made of nine different numbers.
-\item In every column $j$ (i.e., $c_{1,j}$, $c_{2,j}$, ..., $c_{9,j}$) there is at least one element that takes one value in \{1,2,3,4,5,6,7,8,9\}, but all values from 1 to 9 must be taken.
+\item In every column $j$ (i.e., $c_{1,j}$, $c_{2,j}$, $c_{3,j}$, $c_{4,j}$, $c_{5,j}$, $c_{6,j}$, $c_{7,j}$, $c_{8,j}$, $c_{9,j}$) there is at least one element that takes one value in \{1,2,3,4,5,6,7,8,9\}, but all values from 1 to 9 must be taken.
 Since in a column there are only 9 elements, it means that there will be a one-to-one mapping from these column elements to the \{1,2,3,4,5,6,7,8,9\} set values, thus each column will be made of nine different numbers.
@@
 That means that the nine elements along a row, column or within a box must be different and equal to a digit in set $\{1,2,3,4,5,6,7,8,9\}$, thus meaning that a row, column or box is valid if it contains all digits.
-Let's consider $c_1,\,c_2,\,...,\,c_9$, which may be the elements within a row, column or box.
-Then we have:
+Let's consider $c_1,\,c_2,\,c_3,\,c_4,\,c_5,\,c_6,\,c_7,\,c_8,\,c_9$, which may be the elements within a row, column or box. Then we have:
 \begin{align*}
@@
 \wedge &\bigwedge^9_{j=1} \text{ is\_valid}\left( c_{1,j},  c_{2,j},  c_{3,j},  c_{4,j},  c_{5,j},  c_{6,j},  c_{7,j},  c_{8,j},  c_{9,j} \right) \\
-	\wedge &\bigwedge^9_{i,j \in \{1,\,4,\,7\}} \text{ is\_valid}\left( c_{i,1},
+	\wedge &\bigwedge^9_{i,j \in \{1,\,4,\,7\}} \text{ is\_valid}\left( c_{i,j},
 	c_{i,(j+1)},  c_{i,(j+2)}, c_{(i+1),j},  c_{(i+1),(j+1)},
@@
 \begin{align*}
-\text{sudoku}\left( \{ c_{i,j} \}_{i, j \in \{ 1,...,9\}} \right) = & \nonumber\\
+\text{sudoku}\left( \{ c_{i,j} \}_{i, j \in \{ 1,2,3,4,5,6,7,8,9 \}} \right) = & \nonumber\\
        & \bigwedge^9_{i=1} \text{ is\_valid}\left(  c_{i,1},  c_{i,2},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9}  \right)\\
 \wedge &\bigwedge^9_{j=1} \text{ is\_valid}\left( c_{1,j},  c_{2,j},  c_{3,j},  c_{4,j},  c_{5,j},  c_{6,j},  c_{7,j},  c_{8,j},  c_{9,j} \right) \\
-\wedge &\bigwedge^9_{i,j \in B} \text{ is\_valid}\left( c_{i,1},  c_{i,(j+1)},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9} \right).
+\wedge &\bigwedge^9_{i,j \in B} \text{ is\_valid}\left( c_{i,j},  c_{i,(j+1)},  c_{i,(j+2)},  c_{(i+1),j},  c_{(i+1),(j+1)},  c_{(i+1),(j+2)},  c_{(i+2),j},  c_{(i+2),(j+1)},  c_{(i+2),(j+2)} \right).
 \end{align*}
@@
 \text{is\_valid}\left(  c_1,  c_2,  c_3,  c_4,  c_5,  c_6,  c_7,  c_8,  c_9
-	\right) = \bigwedge^9_{i=1}  \bigvee^9_{i=1} \,\,\, c_i
+	\right) \Leftrightarrow \bigwedge^9_{n=1}  \bigvee^9_{i=1} \,\,\, c_i = n
 \end{align*}

### 2021-05-25 solver/app.py
Support differntly sized tables

@@ -1,8 +1,10 @@
+import math
+
 from flask import Flask, jsonify, render_template, request
 from z3 import And, Int, Or, Solver, sat
 
 
-def is_valid(*data: str):
+def is_valid(max_val: int, *data: str):
     """
     Create SAT condition that ensures
     all the entries in the data array are
@@ -10,48 +12,50 @@ def is_valid(*data: str):
     mixed pre-filled and unknown variables)
     """
     x = []
-    for i in range(0, 9):
+    n = len(data)
+    for i in range(0, n):
         if data[i].isnumeric():
             x.append(int(data[i]))
         else:
             x.append(Int(data[i]))
 
     and_args = []
-    for d in range(0, 9):
+    for d in range(0, max_val):
         or_args = []
-        for i in range(0, 9):
+        for i in range(0, n):
             or_args.append(x[i] == d)
         and_args.append(Or(*or_args))
 
     return And(*and_args)
 
 
-def build_sudoku_solver(board):
+def build_sudoku_solver(board, n):
     """
     Return a solver for a
     given sudoku board
     """
     rows_check = []
-    for i in range(0, 9):
-        rows_check.append(is_valid(board[i][0], board[i][1], board[i][2],
-                                   board[i][3], board[i][4], board[i][5],
-                                   board[i][6], board[i][7], board[i][8]))
+    for i in range(0, n):
+        rows_check.append(is_valid(n, *board[i]))
+
     columns_check = []
-    for i in range(0, 9):
-        columns_check.append(is_valid(board[0][i], board[1][i], board[2][i],
-                                      board[3][i], board[4][i], board[5][i],
-                                      board[6][i], board[7][i],
-                                      board[8][i]))
+    for i in range(0, n):
+        args = []
+        for j in range(0, n):
+            args.append(board[j][i])
+        columns_check.append(is_valid(n, *args))
+
     squares_check = []
-    for i in range(0, 9, 3):
-        for j in range(0, 9, 3):
-            squares_check.append(is_valid(board[i][j], board[i][j + 1],
-                                          board[i][j + 2], board[i + 1][j],
-                                          board[i + 1][j + 1],
-                                          board[i + 1][j + 2],
-                                          board[i + 2][j],
-                                          board[i + 2][j + 1],
-                                          board[i + 2][j + 2]))
+    size_box = int(math.sqrt(n))
+    for i in range(0, n, size_box):
[…]

### 2021-05-25 report/paper.tex
Trying to do a 16x16 sudoku: 2:31:16 it's still running. I could have compiled chrome from source in this time

@@
 
+\subsection{Supporting other board configurations}
+
+In order to support different board configurations that are not of size
+$ 9 \times 9 $, we have to make a number of considerations to figure out
+which sizes we can support:
+First, we know that the highest value that a cell may contain is equal to the
+number of cells in each row (or column): in the ``base'' case of the
+$ 9 \times 9 $ grid, in fact the highest value is 9.
+Now, since we have to fit the exact same number of (distinct) elements both
+in a single row, in a single column and in the ``inner box'', we have to
+be able to construct such ``inner box'' that allow us to fill them with a number
+of element equal to the number or cells per row (or column).
+
+Having these constraints at hand, it is clear how the Sudoku game, with the
+(standard) rule set that has been applied through this project is possible only
+when the grid is of size $ n \times n $ where $ n $ is a perfect square number
+(i.e. $ 9 $ is a perfect square number since it's $ 9 = 3^2 $).
+This happens because the ``inner squares'' have to be of size $ \sqrt{n} $
+in order to have $ \sqrt{n} $ of them in each row: this is needed because
+if we have a square with side of length $ \sqrt{n} $, its area is
+$ \sqrt{n}^2 = n $, meaning we can fit in it $ n $ different elements.
+
 \newpage
+
 \section{Instructions for running}
@@
 \centering
+\includegraphics[width=0.8\textwidth]{pics/grid_4.png}
+\caption{A $ 4 \times 4 $ satisfiable board}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.8\textwidth]{pics/grid_16.png}
+\caption{A $ 16 \times 16 $ unsatisfiable board}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
 \includegraphics[width=0.8\textwidth]{pics/hardest_sudoku.png}

### 2021-05-25 solver/app.py

@@ -79,7 +79,7 @@ def app_solve():
         for j in range(0, n):
             if not board[i][j]:
                 # Make a new unknown variable
-                board[i][j] = "x%d%d" % (i, j)
+                board[i][j] = "x%dy%d" % (i, j)
             else:
                 # Pre-filled cell
                 board[i][j] = str(int(board[i][j]) - 1)

### 2021-05-26 report/paper.tex
apology for poor english
when were you when lab2 dies?

i was sat at home doing OM when stefan ring

‘lab2 is kill’

‘no’

@@
 \includegraphics[width=0.8\textwidth]{pics/solved.png}
-\caption{Press the \texttt{Solve} button to find the soluscion}
+\caption{Press the \texttt{Solve} button to find the solution}
 \end{figure}
@@
 \includegraphics[width=0.8\textwidth]{pics/input_check.png}
-\caption{Errors if a number is not in $ \left[ 1, 2,3,4,5,6,7,8, 9 \right] $}
+\caption{Errors if a number is not in $ \left[ 1, \hdots n \right] $ (where $n$ is the side of the board)}
 \end{figure}

### 2021-05-26 solver/app.py

@@ -49,12 +49,10 @@ def build_sudoku_solver(board, n):
     size_box = int(math.sqrt(n))
     for i in range(0, n, size_box):
         for j in range(0, n, size_box):
-            print("SQUARE %s %s" % (int(i / size_box), int(j / size_box)))
             args = []
             for q in range(0, size_box):
                 for w in range(0, size_box):
                     args.append(board[i + q][j + w])
-            print(args)
             squares_check.append(is_valid(n, *args))
 
     s = Solver()
